Incremental Construction of Programs in a Multitask Setting

### 2013-10-23 drafts/nips/nips2013.tex
More edits.

\begin{abstract}
How can machine learning techniques be used to solve problems whose solutions are best represented as computer programs? For example, suppose a researcher wants to design a probabilistic graphical model for a novel domain. Searching the space of probabilistic models automatically is notoriously difficult and is especially difficult when the latent variables are possible. However, when researchers seem able to relatively easily adapt commonly used modeling motifs to new domains. In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce. This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical concepts'' in much the same way that researchers themselves do, by seeing examples in the literature. One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them. In this work, we describe the SEC algorithm, which attempts to learn a distribution over programs by incrementally finding program components that commonly help to solve problems in a given domain, and we show prelimarily that SEC is able to discover the graphical concepts that underlie many of the common graphical model structures.  
\end{abstract}

@@

The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks. Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs. In each iteration, SEC does two things: 
\begin{itemize}
	\item first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given;
	\item second, it compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
\end{itemize}

@@

\subsection{Inference}
We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
The joint distribution factorizes as
\begin{equation}
P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) \prod_i P(e^i_n | G)
\label{eq:joint}
\end{equation}
From equation (\ref{eq:joint}), the EM updates are
\begin{eqnarray}
q(\rho_n) &\propto& P(t_n | \rho_n) \prod_i P(e^i_n | G^{old})\\
\label{eq:qdist}
G^{new} &=& \operatorname{argmin}_G \left( -\ln P(G) -
\sum_n 
E_{q_n}
 \left[ \ln P(\rho_n | G) \right] \right)
 \label{eq:gmax}
\end{eqnarray}
Exact computation of $q(\cdot)$ in equation (\ref{eq:qdist}) would involve marginalizing over the space of all programs.
Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.

@@

\section{Learning to Construct Graphical Models}

@@

In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that might be amenable to synthesis by programs.
We use our algorithm to learn the motifs and design patterns that underly certain popular graphical models.
Concretely, our algorithm induces programs that take as input the visible nodes of an undirected graphical model and introduces new latent nodes and edges.

\subsection{Experiments}

The SEC algorithm was provided a set of related tasks, each of which involved the construction of a program that built undirected graphical models.
Each program synthesis task included a set of input/output test cases; the likelihood of a task, given a program, was defined as the fraction of test cases passed.
Example input/output pairs of the target programs are shown in figure (\ref{fig:tasks}).

The initial library of primitive combinators contains four binary operations upon undirected graphical models.
Each of these operations we introduce, called an \emph{undirected union}, serves to combine two smaller models while also introducing new edges.
%Using this representation, many models may be expressed in a compositional fashion when combined with higher-order combinators taken from functional programming languages, such as \emph{map} and \emph{fold}.
The operation of these four undirected unions is diagrammed in figure (\ref{fig:graphcomb}).
%In this representation of graphs, the vertexes are given unique indexes.
%Some of the undirected unions exploit these indexes by adding edges as a function of index.

\begin{figure}
\begin{minipage}[c][8cm][t]{.4\textwidth}
  \vspace*{\fill}
  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
  \caption{Example input/output pairs of learned programs. Each induced program takes as input a list of visible nodes and produces an undirected graphical model.}
  \label{fig:tasks}
\end{minipage}%
\begin{minipage}[c][8cm][t]{.3\textwidth}
  \vspace*{\fill}
  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
  \caption{The operation of the undirected unions. Union: Computes union of vertexes and edges. UnionC: Adds edges between corresponding vertexes. UnionHT: Adds edge between first (head) vertex and last (tail) vertex. UnionD: Preserves uniqueness of distinctly indexed vertexes.}
  \label{fig:graphcomb}\par\vfill
\end{minipage}
\end{figure}

\subsection{Results}

In future work, could be used as high-level moves in the search for a good structure.
Connect to, eg, Charles Kemp's work.

>>>>

\begin{abstract}
How can machine learning techniques be used to solve problems whose solutions are best represented as computer programs? For example, suppose a researcher wants to design a probabilistic graphical model for a novel domain. Searching the space of probabilistic models automatically is notoriously difficult and is especially difficult when the latent variables are possible. However, when researchers seem able to relatively easily adapt commonly used modeling motifs to new domains. In doing so, they draw on abstractions such as trees, chains, grids and plates to constrain and direct the kinds of models they produce. This suggests that before we ask machine learning algorithms to discover parsimonious models of new domains, we should develop techniques that enable our algorithms to automatically learn these ``graphical conepts'' in much the same way that researchers themselves do, by seeing examples in the literature. One natural way to think of these graphical concepts is as programs that take sets of random variables and produce graphical models that relate them. In this work, we describe the SEC algorithm, which attempts to learn a distribution over programs by incrementally finding program components that commonly help to solve problems in a given domain, and we show prelimarily that SEC is able to discover the graphical concepts that underlie many of the common graphical model structures.  
\end{abstract}

@@

The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks. Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13} (of which it an extension), the SEC algorithm maintains a distribution $\mathcal{D}$ over programs and this distribution helps guide its search to relevant parts of of the space of programs. In each iteration, SEC does two things: 
\begin{itemize}
	\item first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given;
	\item second, it compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.
\end{itemize}

@@

\subsection{Inference}
We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the grammar, $G$, given the tasks, $\{t_n\}$.
The joint distribution factorizes as
\begin{equation}
P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) P(\rho_n | G),
\label{eq:joint}
\end{equation}
where $P(\rho_n | G) = \prod_i P(e^i_n | G)$.
From equation (\ref{eq:joint}), the EM updates are
\begin{eqnarray}
q(\rho_n) &\propto& P(t_n | \rho_n) P(\rho_n | G^{\text{old}})\\
\label{eq:qdist}
G^{\text{new}} &=& \operatorname{argmin}_G \left( -\ln P(G) -
\sum_n 
E_{q_n}
 \left[ \ln P(\rho_n | G) \right] \right)
 \label{eq:gmax}
\end{eqnarray}
Exact computation of these updates requires summing over the infinite space of programs that can be generated from the grammar. To approximate this, we can attempt to evaluate 
Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.

@@

\section{Application to Structure Learning}
Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
Could an algorithm, such as SEC, automatically learn these structures?

\section{Application to Graphical Model Structure Learning}

@@

\subsection{Graph Combinators}

Diagrams illustrating the graph combinators. An example program that makes an Ising model.

\subsection{Learning Graphical Models}

Show the actual results for our test cases: ``density estimation of the space of interesting graphical models.''
In future work, could be used as high-level moves in the search for a good structure.
Connect to, eg, Charles Kemp's work.

\begin{figure}
\begin{minipage}[c][8cm][t]{.4\textwidth}
  \vspace*{\fill}
  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
  \caption{test figure one}
  \label{fig:test1}
\end{minipage}%
\begin{minipage}[c][8cm][t]{.3\textwidth}
  \vspace*{\fill}
  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
  \caption{test figure two}
  \label{fig:test2}\par\vfill
  % \includegraphics[width=5cm,height=4.5cm]{im
  \caption{test figure three}
  \label{fig:test3}
\end{minipage}
\end{figure}
